---
title: "Plot Social Networks"
author: "Miriam Schwyck"
date: "2025-03-25"
output: 
  html_document:
    toc: true
    toc_depth: 6
---

```{r setup, include=FALSE}
# set default options for cells
knitr::opts_chunk$set(echo = TRUE)
# set random seed
set.seed(1234)

# load igraph package, install if necessary
if(!require(igraph)) install.packages("igraph")
# alternatively, if you already have igraph installed, you can just load it
library(igraph)
```

## Read in data

Here, we assume that the data is stored in the same directory as this script.

```{r read_data}
# set your working directory - change this to wherever you saved this script and the data file(s)
wd <- "./"

# read in data in edge list form
edgelist_data <- read.csv(paste0(wd,"social_network_edgelist.csv"))

# look at data
edgelist_data
head(edgelist_data)
View(edgelist_data)
```

## Make and view the network

```{r}
# make the graph object
g <- graph_from_data_frame(edgelist_data, directed=FALSE)
g
```

```{r}
# plot the graph
plot(g, main="Friendship Network")
```

```{r}
# Get the adjacency matrix
get.adjacency(g, sparse=FALSE)

# Get the weighted adjacency matrix
get.adjacency(g, attr="weight", sparse=FALSE)
```


### EXERCISE: Make a directed graph and plot it.

Notice that when we created the graph object, we set the "directed" argument to False.
To create a directed graph (i.e., one where Alice might like Bob but Bob might not like Alice),
change that argument to True.
```{r}
# make a directed graph object and plot it here

```


## Vertices/Nodes

```{r}
# get the vertices/nodes/people
V(g)
```


```{r}
# get number of vertices (i.e., number of people in the network)
n_vertices <- length(V(g))
n_vertices
```

### Get/Set node attributes

Attributes that are stored in the graph object
```{r}
# get the names of the people/nodes/vertices
V(g)$name
```

Add attributes
```{r}
# add age to the graph object
ages <- c(19, 20, 21, 22, 25, 30, 18, 19, 33, 28)
V(g)$age <- ages
print(V(g)$age)
```

Change plot to reflect age
```{r}
# "size" is a special attribute that plot() knows, so you can set it equal to any 
# attribute to visualize that attribute in the plot
V(g)$size <- V(g)$age
# plot again
plot(g)
```

#### EXERCISE: Add another node attribute (e.g., height) and plot the graph that reflects that attribute
```{r}

```


### Calculate node-level social network measures

#### Degree

Degree is the number of edges a node has
```{r}
# get the number of friends each person has (i.e., degree)
degree(g)
hist(degree(g))
```


```{r}
# change the size of the vertex based on that vertex's degree
V(g)$size <- degree(g)*2 + 10
plot(g)
```

In a directed graph, there are two types of degree: 
in-degree (the number of ties pointing into a node)
```{r}
degree(g, mode="in")
```
out-degree (the number of ties pointing out of a node)
```{r}
degree(g, mode="out")
```

#### Strength
Strength is the sum of the weighted edges connected to a node
```{r}
strength(g)
```


#### Eigenvector Centrality

Eigenvector Centrality measures the extent to which a node is connected to other well-connected nodes.
```{r}
eigen_centrality(g)

# notice that it's not just a list of values, so for this one, we have to select the "vector" variable
evc <- eigen_centrality(g)$vector

V(g)$size <- evc*20+10
plot(g)
```


#### Brokerage

Betweenness is the extent to which a node sits on the shortest path between two other nodes.
In other words, how often do you have to go through that node to reach other nodes?
```{r}
betweenness(g)
```

Constraint is an inverse measure of brokerage. It reflects the extent to which a node is
constrained to its local area.
```{r}
constraint(g)
```

##### EXERCISE: Change the node size of the plot to reflect betweenness? 
```{r}

```

##### EXERCISE: Change the node size of the plot to reflect constraint? 
```{r}

```



## Edges

```{r}
# get edges/ties/friendships
E(g)
```

```{r}
# print out each edge ID in a for-loop
for (edge in E(g)) {
  # get edge id
  print(edge)
  # get edge's vertices
  print(E(g)[edge])
}
```

### Get edge attributes

```{r}
E(g)$weight
```


Plot graph based on attributes

```{r}
# make ties in plot thicker based on how strong (either positive or negative) their relationship is
E(g)$width <- abs(E(g)$weight)*5
# plot with weight
plot(g)
```


```{r}
# add color attribute to each edge
for (edge in E(g)) {
  # get edge id
  print(edge)
  # get edge weight
  edge_weight <- E(g)[edge]$weight
  # determine if edge was positive or negative
  # if positive, make the edge green
  # if negative, make the edge red
  if (edge_weight >= 1) {
    edge_color <- "blue"
  } else {
    edge_color <- "darkred"
  }
  E(g)[edge]$color <- edge_color
}
# look at edge attributes
E(g)$weight
E(g)$color

# plot with edge weight and color
plot(g)
```

#### EXERCISE: Add another edge attribute and plot the graph that reflects that attribute
```{r}

```


### Calculate other social network measures
#### Triads
```{r}
# https://r.igraph.org/reference/triad_census.html
triad_census(g) 
triad_count_0 <- triad_census(g)[1]
triad_count_1 <- triad_census(g)[3]
triad_count_2 <- triad_census(g)[11]
triad_count_3 <- triad_census(g)[16]
# get the probability that adjacent vertices of a vertex are connected (i.e., if two nodes share a friend, are they also friends)
transitivity(g)
```

#### Clustering
```{r}
com <- fastgreedy.community(g)
mem <- membership(com)
plot(com, g)
```
